Mapping Mixtures of Genotype-by-Weather Interactions in Switchgrass (Panicum virgatum)

### 2024-02-21 manuscript/revision2/revision2.tex
manuscript: M&M edits and some QTL incorporation

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which allows effects to covary between gardens in many ways
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). To do this, we used
GWAS on site-specific BLUPs for the timing of vegetative growth and
flowering to determine preliminary single nucleotide polymorphism (SNP)
effect estimates for both traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using a subset of randomly selected SNPs used in the GWAS
and their corresponding effect size estimates. In these models
(hereafter referred to as ``random''), we then specified a variety of
potential covariance structures among genetic effect size estimates from
each garden (as expanded on below) and used a greedy algorithm to
iteratively select and add covariance structures that significantly
improved the log likelihood of the \emph{mash} model. Given the high
correlation between some of the covariance patterns we tested, this
approach allowed us to select only the subset of patterns of covariance
that significantly improved the fit of the jointly inferred structure of
genetic effect estimates present in the data. We then selected the set
of SNPs and corresponding effect size estimates that had the largest
effect estimates in univariate GWAS (hereafter ``strong'' effects) and
jointly re-estimated these effect size estimates at all eight sites
using the set of significant covariance matrices that most improved the
model fit on the random effects.

We specified three categories of covariance structures across gardens:
``canonical'' covariance, with simple patterns of effect size covariance
introduced in the initial \emph{mash} manuscript; ``data-driven''
covariances derived from common patterns of SNP effects observed in the
data, and ``GxWeather'' covariance, estimated from the covariance of
empirical weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example,
GxWeather covariance between gardens in MI and MO, for rainfall in the
seven days prior to flowering, captures the covariance in the amount of
rainfall genotypes A, B, and C receive in the seven days prior to their
three flowering dates in MI and their three flowering dates in MO. If
these six rainfall values are similar, covariance will be near one; if
they are uncorrelated, covariance will be near zero; if there is
sufficient rank changing in rainfall amounts between gardens for these
genotypes, this covariance will be near negative one. We specified
multiple covariance structures for the canonical, data-driven, and
GxWeather categories, each of which represented a specific pattern of
covariance in genetic effect size estimates that could be included in
the model.

The GxWeather covariance structures allow hypothesis testing of specific
weather variables as cues for the start of vegetative and reproductive
growth. Say that a SNP in FLC controls flowering in a
photoperiod-dependent manner. In that case, the joint estimate of
effects for that SNP could have a high mixture proportion, or mass, on a
covariance matrix created using a photoperiod-based environmental cue,
such as day length two days prior to flowering. In our data, we would
infer that the effect of that SNP on flowering was caused by a response
to the environmental cue used to construct the GxWeather covariance
structure with the largest mass. Overall, loadings of genetic effects on
these GxWeather matrices in the model, along with the posterior effect
size estimates provide information on genome-wide patterns of
SNP-environment interaction.

The phenotypic correlations for the timing of vegetative growth had
moderate negative correlations between the Texas and North gardens,
particularly in the Gulf subpopulation and Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, then they may be controlled by SNP effects that
differed in sign across these regions (effects with antagonistic
pleiotropy), SNP effects that are large in one garden or region and
non-significant in others (effects with amplification), or a combination
of these patterns. Five GxWeather covariance structures were selected by
the greedy algorithm, one to two per subpopulation. Of these five, three
had mass on them in mash models of the strong effects
(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
covariances between sites between Texas and North gardens
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances. SNP-associated phenotypic effects covaried with different
weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix constructed using the covariance of daylength 14 days
prior to the date of vegetative growth. The covariance matrix for this
weather cue was very similar to the pattern of phenotypic correlation
for the timing of vegetative growth in the Gulf subpopulation
(Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash models of the
timing of vegetative growth in the Midwest subpopulation and Both
subpopulations did not include this GxWeather covariance; the Midwest
had no weight on any GxWeather covariance, while Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). Only the Gulf subpopulation and Both
subpopulations had mass on any GxWeather covariance matrices
(Figure~\ref{fig-covar} C).

@@

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. We used the local
false sign rate (lfsr), an analogue of the lfdr that establishes
confidence in the effect sign, not the effect's difference from zero, to
determine significance. We required lfsr significance (p \textless{}
0.05) in both gardens to include effects. This means that our tests for
antagonistic pleiotropy, or a sign change between conditions, carry an
equal statistical burden to those for effects with the same sign. We
also summarized the pairwise patterns of effect sign and magnitude
within and between the Texas and North regions, within the natural
ranges of the Gulf and Midwest subpopulations (@fig-effects).

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens (@fig-effects A).
78.7\% of pairwise comparisons between North and Texas gardens had a
difference in sign, while only 28.6\% and 0.2\% of North-North or
Texas-Texas comparisons had a difference in sign, respectively
(@fig-effects). The majority of pairwise effects for greenup for the
Midwest (\textgreater55\%) and Both (\textgreater85\%) subpopulations
were the same sign, and effects frequently differed in magnitude between
the MO and OK garden and other gardens (@fig-effects X).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (@fig-effects). More effects differed in
magnitude between the Texas and North regions than within these regions
(42.7\% vs \textless20\%; @fig-effects X). The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign (@fig-effects). Most
differences in sign were between TX1, the southernmost garden, and all
other gardens (@fig-effects B). Similarly, more effects that differed in
magnitude included gardens in the Texas region (52.3-55.9\%), and most
effect pairs in the North region were not distinguishable (91.5\%).

@@

We sought additional experimental support for our joint re-estimates of
SNP effects using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites (@fig-qtl
A,B). We conducted quantitative trait loci (QTL) mapping of flowering as
functions of four environmental cues that we also used as covariance
matrices in mash, and identified eight QTL for flowering date, six QTL
for flowering GDD, ten QTL for flowering day length, and eight QTL
flowering day length change, all of which showed QTL by environment
interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7). All flowering QTL
intervals contained at least one SNP significant in at least one mash
run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
tail of significance, whichever was stricter (SI Appendix, Dataset 8).
We also looked for enrichments of mash SNPs in the 1\% tail of
significance (the `mash 1\% tail') within each QTL interval. At the 5\%
level, five QTL had enrichments of SNPs in the mash 1\% tail. Overall,
there were eleven significant enrichments (p \textless{} 0.05,
hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals.
Our QTL intervals had more enrichments of SNPs in the mash 1\% tail than
were found for all but eleven of these sets of random genomic intervals
(@fig-qtl C, p = 0.011). Thus, we were able to experimentally support
our joint re-estimates of SNP effects with a QTL mapping experiment
using a separate mapping population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (@fig-effects A). This result
supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (22--25), and is
the first experimental work using QTL mapping and GWAS across common
gardens to find antagonistic pleiotropy to be common in small genomic
regions (12, 37, 38).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on a photoperiod cue. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. The genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (@fig-qtl C).

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We know still less about the overwintering
parameters that cause variation in the start of vegetative. Though we
could assign SNP effects to a GxWeather covariance structure in the Gulf
subpopulation, we could not do this for the Midwest subpopulation. More
generally, it is difficult to predict the time scales over which
individuals may integrate environmental cues, particularly in perennial
species which may integrate these cues over longer time scales. If this
integration time itself varies between individuals, we cannot select a
covariance structure that reflects this, though these structures would
likely be highly correlated with GxWeather structures we did include.
Mash offers an opportunity to specify multiple environmental cues and
compete them to explain patterns of genetic effects, allowing us to
detect how important these cues are genome-wide, and how strongly each
cue influences each SNP. This is a key development to further improve
our understanding of genetic variation in GxE.

@@

\caption{\label{fig-effects}Comparisons of pairs of jointly re-estimated
effects across eight gardens, for effects with lfsr \textless{} 0.05.
AHM is considering moving this figure to the supplement and
substantially simplifying the figure in the main text, perhaps just
showing total effects in Texas, the North, and maybe Oklahoma.
Essentially making Table 2 into a new, simplified figure. Effect
patterns for re-estimated SNP effects that are significant at pairs of
sites by the local false sign rate test (lfsr), for (a) greenup date and
(b) flowering date, within and between two genetic subpopulations.
Common gardens are arranged in latitudinal order along the x- and
y-axis. Effects in the first and fourth rows differ in sign at these
pairs of gardens (p \textless{} 0.05, lfsr). Effects in the second,
third, fifth and sixth rows are identical in sign (p \textless{} 0.05,
lfsr). Effects in the second and fifth rows additionally differ in
magnitude by a factor of \textgreater0.4, while effects in the third and
sixth row were not distinguishable by magnitude nor sign of the effect.}

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(35). Mash is a statistical method that allows estimation and comparison
of many effects jointly across many different conditions; it improves on
previous methods by allowing multiple, arbitrary correlations in effect
sizes among conditions (SI Appendix, Section S5). To obtain SNP effect
estimates, we first conducted univariate genome-wide association at each
common garden for green-up and flowering date. We then analyzed SNP
effects for the top 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs per condition using mash, as in (32). Details on
these models can be found in (SI Appendix, Section S6).

@@

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}

\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects across all eight common gardens using the mash model,
which allows effects to covary between gardens in many ways
(Figure~\ref{fig-map} D, SI Appendix, Datasets 1-6). We specified three
categories of covariance structures across gardens: ``canonical''
covariance, with simple patterns of effect size covariance introduced in
the initial \emph{mash} manuscript; ``data-driven'' covariances derived
from common patterns of SNP effects observed in the data, and
``GxWeather'' covariance, estimated from the covariance of empirical
weather patterns at each garden at specific times before the
phenological event (Table 1, SI Appendix, Section S2). For example, the
more similar that the amount of rainfall prior to flowering is for
genotypes A and B between gardens in MI and MO, the nearer these
covariances will be to one. If rainfalls are uncorrelated, covariance
will be near zero; if there is rank changing (e.g.~high rainfall before
flowering of A at MO but not MI, and high rainfall before flowering for
B at MI but not MO), this covariance will be near negative one. We
specified multiple covariance structures for the canonical, data-driven,
and GxWeather categories, each of which represented a specific pattern
of covariance in genetic effect size estimates that could be included in
the model.

The GxWeather covariance structures allow hypothesis testing of specific
weather variables as cues for the start of vegetative and reproductive
growth. Say that a SNP in FLC controls flowering in a
photoperiod-dependent manner. In that case, the joint estimate of
effects for that SNP could have a high mixture proportion, or mass, on a
covariance matrix created using a photoperiod-based environmental cue,
such as day length at some interval prior to flowering. In our data, we
would infer that the effect of that SNP on flowering was caused by a
response to the environmental cue used to construct the GxWeather
covariance structure with the largest mass. Overall, loadings of genetic
effects on these GxWeather matrices in the model, along with the
posterior effect size estimates, provide information on genome-wide
patterns of SNP-environment interaction.

The phenotypic correlations for the timing of vegetative growth had
moderate negative correlations between the Texas and North gardens,
particularly in the Gulf subpopulation and Both subpopulations
(Figure~\ref{fig-map} B). If these phenotypic correlations have a
genetic component, then they may be controlled by SNP effects that
differ in sign across these regions (effects with antagonistic
pleiotropy), SNP effects that are large in one garden or region and
non-significant in others (effects with amplification), or a combination
of these patterns. Five GxWeather covariance structures were selected by
the greedy algorithm, one to two per subpopulation. Of these five, three
had mass on them in mash models of the strong effects
(Figure~\ref{fig-covar} A-B). Two of these three matrices had negative
covariances between sites between Texas and North gardens
(Figure~\ref{fig-covar} A), while one had all positive or near-zero
covariances. SNP-associated phenotypic effects covaried with different
weather-based cues in the Gulf \& in Both subpopulations
(Figure~\ref{fig-covar} B). In total, 65\% of the posterior weight of
strong SNP effects in the mash model of Gulf vegetative growth fell on a
covariance matrix constructed using the covariance of daylength 14 days
prior to the date of vegetative growth. The covariance matrix for this
weather cue was very similar to the pattern of phenotypic correlation
for the timing of vegetative growth in the Gulf subpopulation
(Figure~\ref{fig-map} B; Figure~\ref{fig-covar} A). Mash models of the
timing of vegetative growth in the Midwest subpopulation and Both
subpopulations did not include this GxWeather covariance; the Midwest
had no weight on any GxWeather covariance, while Both subpopulations had
non-zero weights on two additional GxWeather covariance types, average
temperature one day prior to green-up, and the day length change in
seconds in the day prior to green-up (Figure~\ref{fig-covar} B). The
average temperature covariance matrix had negative covariances between
Texas and North gardens, though not as strong as the negative phenotypic
correlations seen in Both subpopulations (Figure~\ref{fig-map} B;
Figure~\ref{fig-covar} A). Only the Gulf subpopulation and Both
subpopulations had mass on any GxWeather covariance matrices
(Figure~\ref{fig-covar} C).

@@

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. We used the local
false sign rate (lfsr), an analogue of the lfdr that establishes
confidence in the effect sign, not the effect's difference from zero, to
determine significance. We required lfsr significance (p \textless{}
0.05) in both gardens to include effects. This means that our tests for
antagonistic pleiotropy, or a sign change between conditions, carry an
equal statistical burden to those for effects with the same sign. We
also summarized the pairwise patterns of effect sign and magnitude
within and between the Texas and North regions, within the natural
ranges of the Gulf and Midwest subpopulations
(Figure~\ref{fig-effects}).

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens
(Figure~\ref{fig-effects} A). 78.7\% of pairwise comparisons between
North and Texas gardens had a difference in sign, while only 28.6\% and
0.2\% of North-North or Texas-Texas comparisons had a difference in
sign, respectively (Figure~\ref{fig-effects}). The majority of pairwise
effects for greenup for the Midwest (\textgreater55\%) and Both
(\textgreater85\%) subpopulations were the same sign, and effects
frequently differed in magnitude between the MO and OK garden and other
gardens (Figure~\ref{fig-effects} X).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Figure~\ref{fig-effects}). More effects
differed in magnitude between the Texas and North regions than within
these regions (42.7\% vs \textless20\%; Figure~\ref{fig-effects} X). The
Midwest population had relatively few significant effects for flowering,
but a large proportion of these differed in sign between Texas and North
regions (42.7\%) or within the North region (65.4\%). Finally, in Both
subpopulations, less than 20\% of pairwise effects differed in sign
(Figure~\ref{fig-effects}). Most differences in sign were between TX1,
the southernmost garden, and all other gardens (Figure~\ref{fig-effects}
B). Similarly, more effects that differed in magnitude included gardens
in the Texas region (52.3-55.9\%), and most effect pairs in the North
region were not distinguishable (91.5\%).

@@

We sought additional experimental support for our joint re-estimates of
SNP effects using an independent pseudo-F2 mapping population created
from Gulf \& Midwest individuals and grown at the same sites
(Figure~\ref{fig-qtl} A,B). We conducted quantitative trait loci (QTL)
mapping of flowering as functions of five environmental cues that we
also used as covariance matrices in mash, and identified eight QTL for
flowering date, eight QTL for flowering as a function of day length
change two days prior, one QTL for the start of vegetative growth, and
two QTL for vegetative growth as a function of daylength change one day
prior, all of which showed QTL by environment interactions (SI Appendix,
Fig. S3). All QTL for flowering overlapped one or more homologs from
rice or \emph{A. thaliana} with functionally validated roles in
flowering (SI Appendix, Dataset 7). All flowering and green-up QTL
intervals contained at least one SNP significant in at least one mash
run at a log10-transformed Bayes Factor \textgreater{} 2, or in the 1\%
tail of significance, whichever was stricter (SI Appendix, Dataset 8).
We also looked for enrichments of mash SNPs in the 1\% tail of
significance (the `mash 1\% tail') within each QTL interval. At the 5\%
level, three QTL had enrichments of SNPs in the mash 1\% tail. Overall,
there were five significant enrichments (p \textless{} 0.05,
hypergeometric test) of SNPs in the mash 1\% tail in the QTL intervals;
however, this number of enrichments was found in many random genomic
intervals also (Figure~\ref{fig-qtl} C). Thus, we were able to
experimentally support some of our re-estimates of SNP effects with a
QTL mapping experiment using a separate mapping population.

@@

Our analysis of the timing of vegetative growth in the Gulf and in Both
subpopulations revealed substantial antagonistic pleiotropy in effects
between the Texas and North gardens (@fig-effects A). This result
supports theoretical models that local adaptation should involve
antagonistic pleiotropy at the level of individual loci (22--25), and is
the first experimental work using GWAS across common gardens to find
antagonistic pleiotropy to be common in small genomic regions (12, 37,
38).

Our analysis of the timing of flowering showed that the Gulf and Midwest
subpopulations have distinct GxWeather: flowering timing in the Midwest
subpopulation has photoperiod-related variation, in that flowering
timing covaries with a day length change signal two days before
flowering occurs. In contrast, the Gulf subpopulation does not have
variation in flowering based on a photoperiod cue. Instead, the Gulf
subpopulation has variation in flowering that covaries with the rainfall
that occurs in the week prior to flowering. Three genomic regions
affecting flowering that we re-estimated across all eight sites were
also supported by QTL from an independent mapping population at these
sites (@fig-qtl A).

Identifying the environmental cues that are predictive of, or even
correlated with, plant phenotypic responses remains a major challenge to
studies interrogating gene action across many natural environments. The
GxWeather photoperiod and cumulative rainfall cues we identify here are
functions of the genotypes measured and capture only a minority of SNP
effects on flowering. We know still less about the overwintering
parameters that cause variation in the start of vegetative growth.
Though we could assign SNP effects to a GxWeather covariance structure
in the Gulf subpopulation, we could not do this for the Midwest
subpopulation. More generally, it is difficult to predict the time
scales over which individuals may integrate environmental cues,
particularly in perennial species which may integrate these cues over
longer time scales. If this integration time itself varies between
individuals, we cannot select a covariance structure that reflects this,
though these structures would likely be highly correlated with GxWeather
structures we did include. Mash offers an opportunity to specify
multiple environmental cues and compete them to explain patterns of
genetic effects, allowing us to detect how important these cues are
genome-wide, and how strongly each cue influences each SNP. This is a
key development to further improve our understanding of genetic
variation in GxE.

@@

\caption{\label{fig-effects}AHM is considering moving this figure to the
supplement and substantially simplifying the figure in the main text,
perhaps just showing total effects in Texas, the North, and maybe
Oklahoma. Comparisons of pairs of jointly re-estimated effects across
eight gardens, for effects with lfsr \textless{} 0.05. Essentially
making Table 2 into a new, simplified figure. Effect patterns for
re-estimated SNP effects that are significant at pairs of sites by the
local false sign rate test (lfsr), for (a) greenup date and (b)
flowering date, within and between two genetic subpopulations. Common
gardens are arranged in latitudinal order along the x- and y-axis.
Effects in the first and fourth rows differ in sign at these pairs of
gardens (p \textless{} 0.05, lfsr). Effects in the second, third, fifth
and sixth rows are identical in sign (p \textless{} 0.05, lfsr). Effects
in the second and fifth rows additionally differ in magnitude by a
factor of \textgreater0.4, while effects in the third and sixth row were
not distinguishable by magnitude nor sign of the effect.}

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(35). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (39). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for both traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
a set of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

@@

\bibitem[\citeproctext]{ref-lovell2021}
\CSLLeftMargin{39. }%
\CSLRightInline{J. T. Lovell, \emph{et al.},
\href{https://doi.org/10.1038/s41586-020-03127-1}{Genomic mechanisms of
climate adaptation in polyploid bioenergy switchgrass}. \emph{Nature},
1--7 (2021).}
